\mode*

% Since this a solution template for a generic talk, very little can
% be said about how it should be structured. However, the talk length
% of between 15min and 45min and the theme suggest that you stick to
% the following rules:  

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.


\section[Bootstrapping]{Bootstrapping authentication}

\subsection{What is bootstrapping?}

\begin{frame}
  \begin{block}{Bootstrapping: A chicken-and-egg problem}
    \begin{itemize}
      \item Alice is not registered in our authentication system.
      \item We want to register her as a user in our system.

        \pause{}

      \item How do we know Alice is actually Alice?
      \item Since she's not registered, we don't have anything to authenticate 
        her.
    \end{itemize}
  \end{block}

  \pause{}

  \begin{exercise}
    \begin{itemize}
      \item Any quick workarounds that comes to mind?
      \item When is this a problem and when is it not?
    \end{itemize}
  \end{exercise}
\end{frame}

\begin{frame}
  \begin{solution}[We don't care who Alice is]
    \begin{itemize}
      \item We simply set up authentication when Alice creates the account.
      \item Now we can authenticate whoever set up the account.
    \end{itemize}
  \end{solution}

  \pause

  \begin{remark}
    \begin{itemize}
      \item The attribute here is account ownership.
    \end{itemize}
  \end{remark}

  \begin{example}
    \begin{itemize}
      \item This is the solution used by most web services.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{solution}[We care who Alice actually is]
    \begin{itemize}
      \item We can require ID checks etc.\ to set up the authentication 
        mechanisms using a helpdesk.

        \pause{}

      \item If we have address etc., then we can send the credentials via mail 
        (be it snailmail or email).

    \end{itemize}
  \end{solution}

  \pause

  \begin{remark}
    \begin{itemize}
      \item The attribute here is identity.
      \item ID checks trust in the issuer.
      \item Sending via mail trusts the mail system.
    \end{itemize}
  \end{remark}
\end{frame}

\begin{frame}
  \begin{exercise}
    \begin{itemize}
      \item How is Alice authenticated when she applies for an ID?
    \end{itemize}
  \end{exercise}
\end{frame}

\begin{frame}
  \begin{example}[Signal, WhatsApp, \dots]
    \begin{itemize}
      \item Relies on ownership of (mobile) phone number.
      \item Send a text message with a code.
    \end{itemize}
  \end{example}

  \pause

  \begin{remark}
    \begin{itemize}
      \item Assumes trustworthy phone infrastructure and operator.
      \item Phone operator can impersonate.
      \item Government can impersonate (forcing phone provider).
    \end{itemize}
  \end{remark}
\end{frame}

\subsection{Problems and recovery}

\begin{frame}
  \begin{remark}
    \begin{itemize}
      \item How to recover from failure?
    \end{itemize}
  \end{remark}

  \begin{example}
    \begin{itemize}
      \item You sign up for a Google account.
      \item You forget your password \dots
    \end{itemize}
  \end{example}

  \begin{example}
    \begin{itemize}
      \item You sign up for a Facebook account.
      \item Someone guesses your password and takes over your account.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{itemize}
    \item Sometimes this is a continuous process.

    \item The same bootstrapping procedure can sometimes also be used for 
      \emph{recovery from failure}.

    \item Make sure the system can handle forgotten, lost or aged 
      authentication means.
  \end{itemize}
\end{frame}

\subsection{Give the problem to someone else}

\begin{frame}
  \begin{idea}
    \begin{itemize}
      \item We could let someone who has solved the problem already do the 
        authentication for us.
    \end{itemize}
  \end{idea}

  \pause

  \begin{remark}
    \begin{itemize}
      \item This make it easier for the user.

      \item This makes the provider a very attractive target.

      \item We must trust them to do it properly \dots
    \end{itemize}
  \end{remark}
\end{frame}

\begin{frame}
  \begin{example}[We don't care who Alice is]
    We can use Google, Facebook \etc.
  \end{example}

  \begin{example}[We care who Alice is]
    We can use \eg BankID.
  \end{example}

  \begin{remark}
    This third party must have done bootstrapping as rigorously as we would 
    have.
  \end{remark}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}[allowframebreaks]
	\small
  \printbibliography
\end{frame}

