\mode*

% Since this a solution template for a generic talk, very little can
% be said about how it should be structured. However, the talk length
% of between 15min and 45min and the theme suggest that you stick to
% the following rules:  

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.


\section{Machine--user authentication}

\subsection{What's the problem?}

\begin{frame}
  \begin{question}
    \begin{itemize}
      \item How can a user tell something legitimate from something 
        illegitimate?
    \end{itemize}
  \end{question}

  \begin{example}
    \begin{itemize}
      \item Payment terminal in the supermarket?
      \item Email from someone?
      \item Web page?
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{definition}[Spoofing/Masquerading]
    \begin{itemize}
      \item Attacker masquerades as authorized.
      \item To a system: impersonates authorized user.
      \item To a user: impersonates legitimate system/UI/part of UI.
    \end{itemize}
  \end{definition}
\end{frame}

\begin{frame}
  \begin{definition}[Phishing]
    \begin{itemize}
      \item A masquerading attack trying to collect sensitive data.
    \end{itemize}
  \end{definition}

  \begin{example}
    \begin{itemize}
      \item Email from IT department requesting the password.
      \item See \url{https://phishingquiz.withgoogle.com/}.
    \end{itemize}
  \end{example}
\end{frame}

\subsection{How to solve?}

\begin{frame}
  \begin{exercise}
    How can we prevent spoofed interfaces?
  \end{exercise}
\end{frame}

\begin{frame}
  \begin{solution}[In general]
    \begin{itemize}
      \item Must have \emph{some} trusted interface.
    \end{itemize}
  \end{solution}

  \begin{example}
    \begin{itemize}
      \item The computer and web browser are trusted.
      \item Web pages are not.
      \item But the browser tries to aid.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{example}
    \begin{itemize}
      \item The computer, browser and web pages are untrusted.
      \item The mobile phone with BankID is trusted.
    \end{itemize}
  \end{example}

  \pause

  \begin{example}
    \begin{itemize}
      \item The payment terminal is untrusted.
      \item The payment card is trusted.
      \item How to communicate with the payment card? --- Through the untrusted 
        terminal.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{example}
    \begin{itemize}
      \item Windows has a trusted path.
      \item Uses the Ctrl+Alt+Del to bring up the authentication dialogue upon 
        login.

        \pause

      \item How many know that?
    \end{itemize}
  \end{example}
\end{frame}

%\begin{frame}
%  \begin{remark}[Problem with social engineering]
%    \begin{itemize}
%      \item These are attacks on higher levels, e.g.\ an email or phone call.
%      \item Difficult to check algorithmically.
%    \end{itemize}
%  \end{remark}
%
%  \begin{example}
%    \begin{itemize}
%      \item Phone call to helpdesk from a \enquote{user} in need.
%      \item Stressful situation, willingness to help, \dots
%    \end{itemize}
%  \end{example}
%
%  \pause{}
%  
%  \begin{example}[Solution?]
%    \begin{itemize}
%      \item Authenticated phone calls.
%      \item E.g.\ display caller ID clearly.
%      \item \enquote{My phone is out of battery, I borrowed a student's}.
%    \end{itemize}
%  \end{example}
%\end{frame}

\begin{frame}
  \begin{block}{Summary}
    \begin{itemize}
      \item Users must have trusted devices.
      \item Users must not make mistakes setting these up.
      \item We must do great usability to aid them.
    \end{itemize}
  \end{block}

  \pause

  \begin{remark}
    \begin{itemize}
      \item This is a very hard problem to solve.
    \end{itemize}
  \end{remark}
\end{frame}


%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}[allowframebreaks]
	\small
  \printbibliography
\end{frame}

