\mode*

% Since this a solution template for a generic talk, very little can
% be said about how it should be structured. However, the talk length
% of between 15min and 45min and the theme suggest that you stick to
% the following rules:  

% - Exactly two or three sections (other than the summary).
% - At *most* three subsections per section.
% - Talk about 30s to 2min per frame. So there should be between about
%   15 and 30 frames, all told.


\section{Something you are}

\subsection{Active and passive biometrics}

\begin{frame}
  \begin{example}[Passive biometrics]
    \begin{itemize}
      \item Fingerprint
      \item Irises
      \item DNA
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{example}[Active biometrics]
    \begin{itemize}
      \item Typing speed
      \item Cursor movement
      \item Web surfing behaviour
      \item Pressure points in signature
    \end{itemize}
  \end{example}
\end{frame}

\subsection{Passive biometrics}

\begin{frame}
  \begin{exercise}
    \begin{itemize}
      \item What about something you are (passive biometrics)?
      \item When is that more than merely a password?
    \end{itemize}
  \end{exercise}
\end{frame}

\begin{frame}
  \begin{example}[Fingerprints for iPhones]
    \begin{itemize}
      \item The iPhone can trust its built-in fingerprint reader.
      \item We know that we read an actual finger.
      \item We know when we read it, so we have freshness.
    \end{itemize}
  \end{example}

  \pause

  \begin{example}[Fingerprints for web services]
    \begin{itemize}
      \item Fingerprint reader connected to laptop.
      \item Browser scans and sends fingerprint to server.
      \item Anyone could send that data, without the reader.
      \item We lack freshness guarantees.
    \end{itemize}
  \end{example}
\end{frame}

\begin{frame}
  \begin{remark}
    \begin{itemize}
      \item We need freshness.
      \item We must store the fingerprint somewhere, to compare.
      \item Someone can copy that data.
      \item Without freshness they can use it --- as a password.
    \end{itemize}
  \end{remark}

  \pause

  \begin{remark}
    \begin{itemize}
      \item Similar for interactive biometrics.
      \item Record and replay.
    \end{itemize}
  \end{remark}
\end{frame}

\begin{frame}
  \begin{exercise}
    \begin{itemize}
      \item How can we ensure freshness?
      \item What data do we need to store for this?
      \item How can we secure that data?
    \end{itemize}
  \end{exercise}
\end{frame}



%%%%%%%%%%%%%%%%%%%%%%


\begin{frame}[allowframebreaks]
	\small
  \printbibliography
\end{frame}

